Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2020-07-21 manuscript/response_to_reviews.Rmd

336 beautifully said “impossible for trees to efficiently transport water to great heights and simultaneously maintain strong resistance and resilience to drought-induced embolism”

**Thank you.**

349 fascinating; it seems plausible that current climate changed induced large-individual mortality involves trees growing to conduit diameters, and thus heights, permitted under previously more even and moister climates, and now finding themselves in more extreme climates that make their heights and conduits excessively wide given current conditions.

**We have added the following text:**

*This may be because moister habitats would tend to support species and individuals with more mesophytic traits (*Bartlett et al. 2016 Drought. DOI: 10.1890/15-0468.1*; *Mencuccini 2003, DOI: 10.1046/j.1365-3040.2003.00991.x*) [@medeiros_extensive_2019], potentially growing to greater heights (e.g., *Detto et al. DOI: 10.1371/journal.pone.0076296*), and these are then more vulnerable when drought hits.*

### 2020-07-21 manuscript/climate_sensitivity_draft.Rmd
sentence on traits and TWI
#122

Belowground, taller trees would tend to have larger root systems (*Enquist and Niklas 2001; DOI: 10.1126/science.1066360*), but this does not necessarily imply that they have greater access to or reliance on deep soil-water resources that may be critical during drought.
Rather, larger trees may allocate more to abundant shallow roots that are beneficial for taking up water from rainstorms (*Meinzer et al. 1999; DOI: 10.1007/s004420050931*).
In any case, the potentially greater access to water did not override the disadvantage conferred by height--and, in fact, greater moisture access in non-drought years (here, higher TWI) appears to make trees more sensitive to drought [@zuleta_drought-induced_2017; @stovall_tree_2019].
The observed height-sensitivity of $Rt$, together with the lack of conferred advantage to large stature in drier topographic positions, agrees with the concept that physiological limitations to transpiration under drought shift from soil water availability to the plant-atmosphere interface as forests age [@bretfeld_plant_2018], such that tall, dominant trees are the most sensitive in mature forests. 
Again, additional research comparing drought responses across forests with different tree heights and water availability would be valuable for disentangling the relative importance of above- and belowground mechanisms across trees fo different size.

>>>>

Belowground, taller trees would tend to have larger root systems (*Enquist and Niklas 2001; DOI: 10.1126/science.1066360*), but this does not necessarily imply that they have greater access to or reliance on deep soil-water resources that may be critical during drought.
Rather, larger trees may allocate more to abundant shallow roots that are beneficial for taking up water from rainstorms (*Meinzer et al. 1999; DOI: 10.1007/s004420050931*).
In any case, the potentially greater access to water did not override the disadvantage conferred by height--and, in fact, greater moisture access in non-drought years (here, higher TWI) appears to make trees more sensitive to drought [@zuleta_drought-induced_2017; @stovall_tree_2019].
This may be because moister habitats would tend to support species and individuals with more mesophytic traits (*Bartlett et al. 2016 Drought. DOI: 10.1890/15-0468.1*; *Mencuccini 2003, DOI: 10.1046/j.1365-3040.2003.00991.x*) [@medeiros_extensive_2019], potentially growing to greater heights (e.g., *Detto et al. DOI: 10.1371/journal.pone.0076296*), and these are then more vulnerable when drought hits.
The observed height-sensitivity of $Rt$, together with the lack of conferred advantage to large stature in drier topographic positions, agrees with the concept that physiological limitations to transpiration under drought shift from soil water availability to the plant-atmosphere interface as forests age [@bretfeld_plant_2018], such that tall, dominant trees are the most sensitive in mature forests. 
Again, additional research comparing drought responses across forests with different tree heights and water availability would be valuable for disentangling the relative importance of above- and belowground mechanisms across trees fo different size.
